Beta Diversity through Time and Space

### 2020-09-09 analyses/beta_diversity.Rmd
update

### Prepare environment
```{r message=FALSE, warning=FALSE}
library(vegan)
library(lubridate)
library(gdm)
library(bipartite)
library(sf)
library(mgcv)
library(mgcViz)
library(tidyverse)
```

### Climate data
```{r, include=FALSE, echo=FALSE}
gdd <- read_csv("../processed_data/climate.csv") %>%
  mutate(year = factor(year),
         yday = yday(date)) %>%
  dplyr::select(site, year, date, yday, gdd.cum) %>%
  semi_join(read_csv("../processed_data/network.csv"), by = "site")
```

@@

### Load network and survey data and generate site.year and site.month.year webs
```{r}
net <- read_csv("../processed_data/network.csv")%>%
  unite(site.year, site, year, sep = "_", remove = FALSE) %>% # create site.year field that can later be split back up
  unite(site.year.day, site, year, dayofyear, sep = "_", remove = FALSE) %>%
  mutate(year = factor(year)) %>% # convert year to factor
  group_by(site, date) %>%
  arrange(site, date)

survey <- read_csv("../processed_data/floral_survey.csv") %>%
  semi_join(net, by = "plant.sp") %>% # Consider only plants visited (anytime anywhere, not specifically for a given site*date)
  dplyr::select(site, date, plant.sp, flower.cover)

site.year_transects <- net %>%
  group_by(site, year) %>%
  summarize(transects = length(unique(date)))

site_transects <- net %>%
  group_by(site) %>%
  summarize(transects = length(unique(date)))

web_site.year <- net %>%
  group_by(site.year, bb.sp, plant.sp) %>%
  summarize(freq = n()) %>%
  dplyr::select(higher = bb.sp, lower = plant.sp, webID = site.year, freq) %>% # rename columns to match bipartite's expectations
  data.frame() %>% # convert to data frame
  frame2webs() # convert to to bipartite web

web_site <- net %>%
  group_by(site, bb.sp, plant.sp) %>%
  summarize(freq = n()) %>% 
  dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
  data.frame() %>% # convert to data frame
  frame2webs() # convert to to bipartite web

web_site.date.2010 <- net %>%
  filter(year == 2010) %>%
  group_by(site.year.day, bb.sp, plant.sp) %>%
  summarize(freq = n()) %>%
  dplyr::select(higher = bb.sp, lower = plant.sp, webID = site.year.day, freq) %>% # rename columns to match bipartite's expectations
  data.frame() %>% # convert to data frame
  frame2webs() # convert to to bipartite web

web_site.date.2011 <- net %>%
  filter(year == 2011) %>%
  group_by(site.year.day, bb.sp, plant.sp) %>%
  summarize(freq = n()) %>%
  dplyr::select(higher = bb.sp, lower = plant.sp, webID = site.year.day, freq) %>% # rename columns to match bipartite's expectations
  data.frame() %>% # convert to data frame
  frame2webs() # convert to to bipartite web

web_site.date.2012 <- net %>%
  filter(year == 2012) %>%
  group_by(site.year.day, bb.sp, plant.sp) %>%
  summarize(freq = n()) %>%
  dplyr::select(higher = bb.sp, lower = plant.sp, webID = site.year.day, freq) %>% # rename columns to match bipartite's expectations
  data.frame() %>% # convert to data frame
  frame2webs() # convert to to bipartite web

array_site <- webs2array(web_site)
array_site.year <- webs2array(web_site.year)
array_site.date.2010 <- webs2array(web_site.date.2010)
array_site.date.2011 <- webs2array(web_site.date.2011)
array_site.date.2012 <- webs2array(web_site.date.2012)
```

### Tabulate webs
S	= beta_S, the dissimilarity in species composition
OS = beta_OS, the dissimilarity (component) explained by "rewiring" among shared species
WN = beta_WN, the dissimilarity between the two networks
ST = beta_ST, the dissimilarity (component) explained by difference in species community composition 
```{r}
beta_site <- betalinkr_multi(array_site, partition.st = TRUE, partition.rr = TRUE) %>%
  as_tibble() %>%
  rename(site1 = i) %>%
  rename(site2 = j)%>%
  left_join(site_data, by = c("site1" = "site")) %>%
  dplyr::select(site1, site2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1 = elev.mean, man1 = management, lon1 = lon, lat1 = lat) %>%
  left_join(site_data, by = c("site2" = "site")) %>%
  dplyr::select(site1, site2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
  left_join(site_transects, by = c("site1" = "site")) %>%
  dplyr::select(site1, site2, S, OS, WN, ST, 
                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
                elev1, elev2, man1, man2, lon1, lat1, lon2, lat2, transects1 = transects) %>%
  left_join(site_transects, by = c("site2" = "site")) %>%
  dplyr::select(site1, site2, S, OS, WN, ST, 
                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
                elev1, elev2, man1, man2, lon1, lat1, lon2, lat2, transects1, transects2 = transects) %>%
  mutate(elev.diff = abs(elev1 - elev2),
         mean.elev = (elev1 + elev2)/2,
         transects.diff = abs(transects1 - transects2),
         man.diff = case_when(
           man1 == man2 ~ FALSE,
           man1 != man2 ~ TRUE
         )) %>%
  gather(metric, value, -c(site1, site2, elev1, elev2, elev.diff, 
                           man1, man2, man.diff, mean.elev, lon1, 
                           lat1, lon2, lat2, transects1, transects2, transects.diff)) %>%
  mutate(metric.class = case_when(
    metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
    metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
    metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
    metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"
  ))

#write_csv(beta_site, "../analyses/output/beta_site.csv")

beta_site.year <- betalinkr_multi(array_site.year, partition.st = TRUE, partition.rr = TRUE, binary = TRUE) %>%
  as_tibble() %>%
  separate(i, into = c("site1", "year1"), sep = "_") %>%
  separate(j, into = c("site2", "year2"), sep = "_") %>%
  left_join(site_data, by = c("site1" = "site")) %>%
  dplyr::select(site1, year1, site2, year2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1 = elev.mean, man1 = management, lon1 = lon, lat1 = lat) %>%
  left_join(site_data, by = c("site2" = "site")) %>%
  dplyr::select(site1, year1, site2, year2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
  left_join(site.year_transects, by = c("site1" = "site", "year1" = "year")) %>%
  dplyr::select(site1, site2, year1, year2, S, OS, WN, ST, 
                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
                elev1, elev2, man1, man2, lon1, lat1, lon2, lat2, transects1 = transects) %>%
  left_join(site.year_transects, by = c("site2" = "site", "year2" = "year")) %>%
  dplyr::select(site1, site2, year1, year2, S, OS, WN, ST, 
                ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
                elev1, elev2, man1, man2, lon1, lat1, lon2, lat2, transects1, transects2 = transects) %>%
  mutate(elev.diff = abs(elev1 - elev2),
         mean.elev = (elev1 + elev2)/2,
         transects.diff = abs(transects1 - transects2),
         man.diff = case_when(
           man1 == man2 ~ FALSE,
           man1 != man2 ~ TRUE
         )) %>%
  gather(metric, value, -c(site1, year1, site2, year2, elev1, elev2, elev.diff, 
                           man1, man2, man.diff, mean.elev, lon1, lat1, lon2, lat2,
                           transects1, transects2, transects.diff)) %>%
  mutate(metric.class = case_when(
    metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
    metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
    metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
    metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"
  ))

### By individual transect

beta_site.date.2010 <- betalinkr_multi(array_site.date.2010, 
                                       partition.st = TRUE, 
                                       partition.rr = TRUE) %>%
  as_tibble() %>%
  separate(i, into = c("site1", "year1", "yday1"), sep = "_") %>%
  separate(j, into = c("site2", "year2", "yday2"), sep = "_") %>%
  left_join(site_data, by = c("site1" = "site"))%>%
  dplyr::select(site1, year1, yday1, site2, year2, yday2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1 = elev.mean, man1 = management, lon1 = lon, lat1 = lat) %>%
  left_join(site_data, by = c("site2" = "site")) %>%
  dplyr::select(site1, year1, yday1, site2, year2, yday2, S, OS, WN, ST, 
         ST.l, ST.h, ST.lh, WN.repl, OS.repl, WN.rich, OS.rich, 
         elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
  mutate(site1 = factor(site1),
         site2 = factor(site2),
         year1 = factor(year1),
         year2 = factor(year2),
         yday1 = as.numeric(yday1),
         yday2 = as.numeric(yday2),
         elev.diff = abs(elev1 - elev2),
         mean.elev = (elev1 + elev2)/2,
         yday.diff = abs(yday1 - yday2),
         man.diff = case_when(
           man1 == man2 ~ FALSE,
           man1 != man2 ~ TRUE
         )) %>%
  gather(metric, value, -c(site1, year1, yday1, site2, year2, yday2,
                           elev1, elev2, elev.diff, yday.diff, 
                           man1, man2, man.diff, mean.elev, 
                           lon1, lat1, lon2, lat2)) %>%
  mutate(metric.class = case_when(
    metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
    metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
    metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
    metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"
  ))

write_csv(beta_site.date, "../analyses/output/beta_site.date.csv")
```

### Visualize transect-level beta-diversity
```{r}
ggplot(filter(beta_site.date.2010, metric != "S" & 
                metric.class != "OS.partition"), 
       aes(elev.diff, value, color = metric)) +
  geom_point(alpha = 0.05) +
  geom_smooth() +
  facet_wrap(~metric.class)

ggplot(filter(beta_site.date.2010, yday.diff <= 7 &
                metric != "S" & 
                metric.class != "OS.partition"), 
       aes(elev.diff, value, color = metric)) +
  geom_point(alpha = 0.05) +
  geom_smooth() +
  facet_wrap(~metric.class)

ggplot(filter(beta_site.date.2011, yday.diff <= 7 &
                metric != "S" & 
                metric.class != "OS.partition"), 
       aes(elev.diff, value, color = metric)) +
  geom_point(alpha = 0.05) +
  geom_smooth() +
  facet_wrap(~metric.class)

ggplot(filter(beta_site.date.2012, yday.diff <= 7 &
                metric != "S" & 
                metric.class != "OS.partition"), 
       aes(elev.diff, value, color = metric)) +
  geom_point(alpha = 0.05) +
  geom_smooth() +
  facet_wrap(~metric.class)

ggplot(filter(beta_site.date.2012, yday.diff == 0 &
                metric != "S" & 
                metric.class != "OS.partition"),
       aes(value)) +
  geom_histogram()  +
  facet_wrap(~metric)

ggplot(filter(beta_site.date.2012, yday.diff <= 10 &
                metric != "S" & 
                metric.class != "OS.partition"),
       aes(value)) +
  geom_histogram()  +
  facet_wrap(~metric)

ggplot(filter(beta_site.date.2012, yday.diff <= 7 &
                metric != "S" & 
                metric.class != "OS.partition" &
                site1 == site2),
       aes(value)) +
  geom_histogram()  +
  facet_wrap(~metric)

```

# Beta diversity by elevation difference
```{r}
# Interaction beta diversity by elevation difference (by year)
## This shows that the patterns are extremely consistent across years, which is not surprising, since they are driven overwhelmingly by floral turnover, and plants tend to be in the same place from year to year.
ggplot(filter(beta_site.year, year1 == year2 & 
                metric != "S" & 
                metric.class != "OS.partition"), 
       aes(elev.diff, value, color = metric)) +
  geom_point(alpha = 0.25) +
  geom_smooth() +
  facet_grid(metric.class ~ year1)

ggplot(filter(beta_site.year, year1 == year2 & 
                metric != "S" & 
                metric.class != "OS.partition"), 
       aes(transects.diff, value, color = metric)) +
  geom_point(alpha = 0.25) +
  geom_smooth() +
  facet_grid(metric.class ~ year1, scales = "free_x")

# Interaction beta diversity by elevation difference, this time pooling across years
ggplot(filter(beta_site, metric != "S" & 
                metric.class != "OS.partition"), 
       aes(elev.diff, value, color = metric)) +
  geom_point(alpha = 0.05) +
  geom_smooth() +
  facet_wrap(~metric.class)

# How important is the difference in transects? Not as important as the difference in elevation.
ggplot(filter(beta_site, metric != "S" & 
                metric.class != "OS.partition"), 
       aes(transects.diff, value, color = metric)) +
  geom_point(alpha = 0.05) +
  geom_smooth() +
  facet_wrap(~metric.class)

ggplot(beta_site, 
       aes(elev.diff, transects.diff)) +
  geom_point(alpha = 0.05) +
  geom_smooth()

### Okay, we'll just do partial mantel tests to partial out the effect of transects.diff; and we'll only do that for the aggregate measures because it seems a bit silly to doi it for the partitions  
WN_diff <- beta_site %>%
  filter(metric == "WN") %>%
  dplyr::select(site1, site2, value) %>%
  spread(site2, value) %>%
  dist()

WN_diff <- beta_site %>%
  filter(metric == "WN") %>%
  dplyr::select(site1, site2, value) %>%
  spread(site2, value) %>%
  dist()

ST_diff <- beta_site %>%
  filter(metric == "ST") %>%
  dplyr::select(site1, site2, value) %>%
  spread(site2, value) %>%
  dist()

OS_diff <- beta_site %>%
  filter(metric == "OS") %>%
  dplyr::select(site1, site2, value) %>%
  spread(site2, value) %>%
  dist()

elev_diff <- beta_site %>%
  dplyr::select(site1, site2, elev.diff) %>%
  unique() %>%
  spread(site2, elev.diff) %>%
  dist()

transects_diff <- beta_site %>%
  dplyr::select(site1, site2, transects.diff) %>%
  unique() %>%
  spread(site2, transects.diff) %>%
  dist()

WN_mantel <- mantel.partial(elev_diff, WN_diff, transects_diff)
ST_mantel <- mantel.partial(elev_diff, ST_diff, transects_diff)
OS_mantel <- mantel.partial(elev_diff, OS_diff, transects_diff)
```

### Let's look at floral survey diversity patterns
```{r}
### 2012
survey_bioData.2012 <- survey %>%
  mutate(year = year(date),
         yday = yday(date)) %>%
  filter(flower.cover > 0 & year == "2012") %>%
  mutate(plant.sp = str_replace_all(plant.sp, " ", "_")) %>%
  group_by(site, yday, plant.sp) %>%
  summarize(cover = sum(flower.cover)) %>%
  spread(plant.sp, cover) %>%
  replace(is.na(.), 0) %>%
  arrange(site, yday) %>%
  unite(site.day, site, yday, sep = "_")

predData.2012 <- survey %>%
  mutate(year = year(date),
         yday = yday(date)) %>%
  filter(flower.cover > 0 & year == "2012") %>%
  select(site, yday) %>%
  left_join(site_data) %>%
  ungroup() %>%
  dplyr::select(site, elev.mean, lat, lon, yday) %>%
  unique() %>%
  arrange(site, yday) %>%
  unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
  select(-site)

survey_sitepair.2012 <- formatsitepair(survey_bioData.2012, bioFormat = 1, siteColumn = "site.day", dist = "jaccard",
                              XColumn = "lon", YColumn = "lat", predData = predData.2012)

### Call gdm()
gdm_survey.2012 <- gdm(survey_sitepair.2012, geo = TRUE)
#gdm.varImp(survey_sitepair, geo = TRUE)

summary(gdm_survey.2012)
plot(gdm_survey.2012)
#plotUncertainty(survey_sitepair, sampleSites = 0.7, bsIters = 50, geo  = TRUE)

@@

### And now Bombus diversity patterns
```{r}
BB_bioData <- net %>%
  group_by(site, date, bb.sp) %>%
  summarize(count = n()) %>%
  group_by(site, date) %>%
  mutate(prop = count/sum(count)) %>%
  group_by(site, bb.sp) %>%
  summarize(abund = sum(prop)) %>%
  spread(bb.sp, abund) %>%
  replace(is.na(.), 0)

predData <- site_data %>%
  ungroup() %>%
  dplyr::select(site, elev.mean, lat, lon) %>%
  unique()

bb_sitepair <- formatsitepair(BB_bioData, bioFormat = 1, siteColumn = "site", dist = "jaccard",
                              XColumn = "lon", YColumn = "lat", predData = predData)

### Call gdm()
gdm_bb <- gdm(bb_sitepair, geo = TRUE)
gdm.varImp(bb_sitepair, geo = TRUE)

summary(gdm_bb)
plot(gdm_bb)
plotUncertainty(bb_sitepair, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
```

### Now let's look at interaction beta diversity
```{r}
net_sitepair <- beta_site %>%
  filter(metric == "WN") %>%
  mutate(weights = rep(1, n())) %>%
  dplyr::select(distance = value, weights,
         s1.xCoord = lon1, s1.yCoord = lat1, 
         s2.xCoord = lon2, s2.yCoord = lat2, 
         s1.elev.mean = elev1, s2.elev.mean = elev2) %>%
  data.frame()

### Call gdm()
gdm_net <- gdm(net_sitepair, geo = TRUE)
gdm.varImp(net_sitepair, geo = TRUE)

summary(gdm_net)
plot(gdm_net)
plotUncertainty(net_sitepair, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
```

### At the transect level
```{r}
beta_site.date_gdm.prep

transects_2010 <- beta_site.date.2010 %>%
  filter(metric == "WN") %>%
  mutate(weights = rep(1, n())) %>%
  dplyr::select(distance = value, weights,
         s1.xCoord = lon1, s1.yCoord = lat1, 
         s2.xCoord = lon2, s2.yCoord = lat2, 
         s1.elev = elev1, s2.elev = elev2,
         s1.yday = yday1, s2.yday = yday2) %>%
  data.frame()

### Call gdm()
gdm_net <- gdm(transects_2010, geo = TRUE)
gdm.varImp(transects_2010, geo = TRUE)

summary(gdm_net)
plot(gdm_net)
plotUncertainty(net_sitepair, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
```

>>>>

### Prepare environment
```{r message=FALSE, warning=FALSE}
library(lubridate)
library(sf)
library(gdm)
library(bipartite)
library(vegan)
library(mgcv)
library(mgcViz)
library(see)
library(tidyverse)
```

### Climate data
```{r echo=FALSE, message=FALSE, include=FALSE}
gdd <- read_csv("../processed_data/climate.csv") %>%
  mutate(year = factor(year),
         yday = yday(date)) %>%
  dplyr::select(site, year, date, yday, gdd.cum) %>%
  semi_join(read_csv("../processed_data/network.csv"), by = "site")
```

@@

### Load network and survey data
```{r}
net <- read_csv("../processed_data/network.csv")%>%
  unite(site.year, site, year, sep = "_", remove = FALSE) %>% # create site.year field that can later be split back up
  unite(site.year.day, site, year, dayofyear, sep = "_", remove = FALSE) %>%
  mutate(year = factor(year)) %>% # convert year to factor
  group_by(site, date) %>%
  arrange(site, date)

survey <- read_csv("../processed_data/floral_survey.csv") %>%
  semi_join(net, by = "plant.sp") %>% # Consider only plants visited (anytime anywhere, not specifically for a given site*date)
  dplyr::select(site, date, plant.sp, flower.cover)
```

### betaboot function: This bootstraps the partitioning of interaction beta-diversity to account for differences in the number of sampling dates for each site in the same way the my other bootstrapping function ("bootnet" or something like that) bootstraps the calculation of network metrics. Note that due to some annoying behavior by webs2array(), I have to define an object within the for loop  of my function using the <<- global assignment operator. This is bad coding, but I can't figure out another way to get webs2array() to find the object it needs to find.
```{r message=FALSE}
betaboot <- function(x, n, iter) {

  out <- data.frame()

  index <- 0

  t_filter <- x %>%
    dplyr::select(site, date, year) %>%
    unique() %>%
    group_by(site, year) %>%
    mutate(transects = n()) %>%
    filter(transects >= n)

  for(i in 1:iter){

    index <- index + 1

    net.sub <- x %>%
      semi_join(t_filter) %>%
      ungroup() %>%
      dplyr::select(site, dayofyear) %>%
      unique() %>%
      group_by(site) %>%
      sample_n(n) %>%
      left_join(x)

    web <<- net.sub %>%
      group_by(site, bb.sp, plant.sp) %>%
      summarize(freq = n()) %>%
      dplyr::select(higher = bb.sp, lower = plant.sp, webID = site, freq) %>% # rename columns to match bipartite's expectations
      data.frame() %>% # convert to data frame
      frame2webs() # convert to to bipartite web

    array <- webs2array(web)

    beta <- betalinkr_multi(array, partition.st = TRUE, partition.rr = TRUE, binary = TRUE) %>%
      as_tibble() %>%
      mutate(iteration = rep(index, n()))

    out <- bind_rows(out, beta)
  }

  out <- out %>%
    as_tibble() %>%
    dplyr::select(site1 = i, site2 = j, everything()) %>%
    gather(metric, value, -c(site1, site2, iteration)) %>%
    mutate(metric.class = case_when(
        metric %in% c("WN", "ST", "OS", "S") ~ "aggregate",
        metric %in% c("WN.repl", "WN.rich") ~ "WN.partition",
        metric %in% c("ST.h", "ST.l", "ST.lh") ~ "ST.partition",
        metric %in% c("OS.repl", "OS.rich") ~ "OS.partition"
      ),
      iteration = factor(iteration))
}
```

### Tabulate webs
S	= beta_S, the dissimilarity in species composition
OS = beta_OS, the dissimilarity (component) explained by "rewiring" among shared species
WN = beta_WN, the dissimilarity between the two networks
ST = beta_ST, the dissimilarity (component) explained by difference in species community composition 
```{r message=FALSE}
net2010 <- filter(net, year == 2010)
net2011 <- filter(net, year == 2011)
net2012 <- filter(net, year == 2012)

beta_2010 <- betaboot(net2010, 5, 50) %>%
  group_by(site1, site2, metric, metric.class) %>%
  summarize(value = mean(value)) %>%
  left_join(site_data, by = c("site1" = "site")) %>%
  dplyr::select(site1, site2, metric, metric.class, value, 
                elev1 = elev.mean, man1 = management, lon1 = lon, lat1 = lat) %>%
  left_join(site_data, by = c("site2" = "site")) %>%
  dplyr::select(site1, site2, metric, metric.class, value, 
         elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
  mutate(elev.diff = abs(elev1 - elev2),
         mean.elev = (elev1 + elev2)/2,
         man.diff = case_when(
           man1 == man2 ~ FALSE,
           man1 != man2 ~ TRUE
         ),
         year = factor(rep("2010", n())))

beta_2011 <- betaboot(net2011, 5, 50) %>%
  group_by(site1, site2, metric, metric.class) %>%
  summarize(value = mean(value)) %>%
  left_join(site_data, by = c("site1" = "site")) %>%
  dplyr::select(site1, site2, metric, metric.class, value, 
                elev1 = elev.mean, man1 = management, lon1 = lon, lat1 = lat) %>%
  left_join(site_data, by = c("site2" = "site")) %>%
  dplyr::select(site1, site2, metric, metric.class, value, 
         elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
  mutate(elev.diff = abs(elev1 - elev2),
         mean.elev = (elev1 + elev2)/2,
         man.diff = case_when(
           man1 == man2 ~ FALSE,
           man1 != man2 ~ TRUE
         ),
         year = factor(rep("2011", n())))

beta_2012 <- betaboot(net2012, 9, 50) %>%
  group_by(site1, site2, metric, metric.class) %>%
  summarize(value = mean(value)) %>%
  left_join(site_data, by = c("site1" = "site")) %>%
  dplyr::select(site1, site2, metric, metric.class, value, 
                elev1 = elev.mean, man1 = management, lon1 = lon, lat1 = lat) %>%
  left_join(site_data, by = c("site2" = "site")) %>%
  dplyr::select(site1, site2, metric, metric.class, value, 
         elev1, elev2 = elev.mean, man1, man2 = management, lon1, lat1, lon2 = lon, lat2 = lat) %>%
  mutate(elev.diff = abs(elev1 - elev2),
         mean.elev = (elev1 + elev2)/2,
         man.diff = case_when(
           man1 == man2 ~ FALSE,
           man1 != man2 ~ TRUE
         ),
         year = factor(rep("2012", n())))

beta_agg <- bind_rows(beta_2010, beta_2011, beta_2012)

#write_csv(beta_agg, "./output/aggregated_beta_diversity.csv")
#beta_agg <- read_csv("./output/aggregated_beta_diversity.csv")
```

### Plot results. There's nothing a Mantel test can tell me that I don't already know from the combination of plotting and GDM analysis.
```{r}
### Elevation
ggplot(filter(beta, year == "2010" & 
                metric != "S" & 
                metric.class != "OS.partition"), 
       aes(elev.diff, value, color = metric)) +
  geom_point(alpha = 0.45) +
  geom_smooth(method = "lm", formula =  y ~ poly(x, 3)) +
  facet_wrap(~metric.class) +
  theme_abyss()

ggplot(filter(beta, year == "2011" & 
                metric != "S" & 
                metric.class != "OS.partition"), 
       aes(elev.diff, value, color = metric)) +
  geom_point(alpha = 0.45) +
  geom_smooth(method = "lm", formula =  y ~ poly(x, 3)) +
  facet_wrap(~metric.class) +
  theme_abyss()

ggplot(filter(beta, year == "2012" & 
                metric != "S" & 
                metric.class != "OS.partition"), 
       aes(elev.diff, value, color = metric)) +
  geom_point(alpha = 0.45) +
  geom_smooth(method = "lm", formula =  y ~ poly(x, 3)) +
  facet_wrap(~metric.class) +
  theme_abyss()

ggplot(filter(beta, 
                metric != "S" & 
                metric.class != "OS.partition"), 
       aes(elev.diff, value, color = metric)) +
  geom_point(alpha = 0.45) +
  geom_smooth(method = "lm", formula =  y ~ poly(x, 3)) +
  facet_wrap(~metric.class) +
  theme_gray()

### Management
ggplot(filter(beta, year == "2010" & 
                metric != "S" & 
                metric.class != "OS.partition"), 
       aes(man.diff, value)) +
  geom_boxplot() +
  facet_wrap(~metric, scales = "free") +
  theme_gray()

ggplot(filter(beta, year == "2011" & 
                metric != "S" & 
                metric.class != "OS.partition"), 
       aes(man.diff, value)) +
  geom_boxplot() +
  facet_wrap(~metric, scales = "free") +
  theme_gray()

ggplot(filter(beta, year == "2012" & 
                metric != "S" & 
                metric.class != "OS.partition"), 
       aes(man.diff, value)) +
  geom_boxplot() +
  facet_wrap(~metric, scales = "free") +
  theme_gray()
```

### Let's look at per-transect floral survey beta-diversity
```{r}
### 2012
survey_bioData.2012 <- survey %>%
  mutate(year = year(date),
         yday = yday(date)) %>%
  filter(flower.cover > 0 & year == "2012") %>%
  mutate(plant.sp = str_replace_all(plant.sp, " ", "_")) %>%
  group_by(site, yday, plant.sp) %>%
  summarize(cover = sum(flower.cover)) %>%
  spread(plant.sp, cover) %>%
  replace(is.na(.), 0) %>%
  arrange(site, yday) %>%
  unite(site.day, site, yday, sep = "_")

survey_predData.2012 <- survey %>%
  mutate(year = year(date),
         yday = yday(date)) %>%
  filter(flower.cover > 0 & year == "2012") %>%
  dplyr::select(site, yday) %>%
  left_join(site_data) %>%
  ungroup() %>%
  dplyr::select(site, elev.mean, lat, lon, yday) %>%
  unique() %>%
  arrange(site, yday) %>%
  unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
  dplyr::select(-site)

survey_sitepair.2012 <- formatsitepair(survey_bioData.2012, bioFormat = 1, siteColumn = "site.day", dist = "jaccard",
                              XColumn = "lon", YColumn = "lat", predData = survey_predData.2012)

### Call gdm()
gdm_survey.2012 <- gdm(survey_sitepair.2012, geo = TRUE)
gdm.varImp(survey_sitepair.2012, geo = TRUE)

summary(gdm_survey.2012)
plot(gdm_survey.2012)
plotUncertainty(survey_sitepair.2012, sampleSites = 0.7, bsIters = 10, geo  = TRUE)

@@

### And now Bombus diversity patterns
```{r}
### 2010
BB_bioData.2010 <- net %>%
  filter(year == 2010) %>%
  group_by(site, date, bb.sp) %>%
  summarize(count = n()) %>%
  group_by(site, date) %>%
  #mutate(prop = count/sum(count)) %>%
  #group_by(site, bb.sp) %>%
  # summarize(abund = sum(prop)) %>%
  spread(bb.sp, count) %>%
  replace(is.na(.), 0) %>%
  mutate(yday = yday(date)) %>%
  unite(site.day, site, yday, sep = "_", remove = "TRUE") %>%
  ungroup() %>%
  select(-date)

BB_predData.2010 <- net %>%
  filter(year == 2010) %>%
  mutate(yday = yday(date)) %>%
  dplyr::select(site, yday) %>%
  left_join(site_data) %>%
  ungroup() %>%
  dplyr::select(site, elev.mean, lat, lon, yday) %>%
  unique() %>%
  arrange(site, yday) %>%
  unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
  dplyr::select(-site)

bb_sitepair.2010 <- formatsitepair(BB_bioData.2010, bioFormat = 1, siteColumn = "site.day", dist = "jaccard",
                              XColumn = "lon", YColumn = "lat", predData = BB_predData.2010)

### Call gdm()
gdm_bb.2010 <- gdm(bb_sitepair.2010, geo = TRUE)
gdm.varImp(bb_sitepair.2010, geo = TRUE)

summary(gdm_bb.2010)
plot(gdm_bb.2010)
plotUncertainty(bb_sitepair.2010, sampleSites = 0.7, bsIters = 50, geo  = TRUE)

### 2011
BB_bioData.2011 <- net %>%
  filter(year == 2011) %>%
  group_by(site, date, bb.sp) %>%
  summarize(count = n()) %>%
  group_by(site, date) %>%
  #mutate(prop = count/sum(count)) %>%
  #group_by(site, bb.sp) %>%
  # summarize(abund = sum(prop)) %>%
  spread(bb.sp, count) %>%
  replace(is.na(.), 0) %>%
  mutate(yday = yday(date)) %>%
  unite(site.day, site, yday, sep = "_", remove = "TRUE") %>%
  ungroup() %>%
  select(-date)

BB_predData.2011 <- net %>%
  filter(year == 2011) %>%
  mutate(yday = yday(date)) %>%
  dplyr::select(site, yday) %>%
  left_join(site_data) %>%
  ungroup() %>%
  dplyr::select(site, elev.mean, lat, lon, yday) %>%
  unique() %>%
  arrange(site, yday) %>%
  unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
  dplyr::select(-site)

bb_sitepair.2011 <- formatsitepair(BB_bioData.2011, bioFormat = 1, siteColumn = "site.day", dist = "jaccard",
                              XColumn = "lon", YColumn = "lat", predData = BB_predData.2011)

### Call gdm()
gdm_bb.2011 <- gdm(bb_sitepair.2011, geo = TRUE)
gdm.varImp(bb_sitepair.2011, geo = TRUE)

summary(gdm_bb.2011)
plot(gdm_bb.2011)
plotUncertainty(bb_sitepair.2011, sampleSites = 0.7, bsIters = 50, geo  = TRUE)

### 2012
BB_bioData.2012 <- net %>%
  filter(year == 2012) %>%
  group_by(site, date, bb.sp) %>%
  summarize(count = n()) %>%
  group_by(site, date) %>%
  #mutate(prop = count/sum(count)) %>%
  #group_by(site, bb.sp) %>%
  # summarize(abund = sum(prop)) %>%
  spread(bb.sp, count) %>%
  replace(is.na(.), 0) %>%
  mutate(yday = yday(date)) %>%
  unite(site.day, site, yday, sep = "_", remove = "TRUE") %>%
  ungroup() %>%
  select(-date)

BB_predData.2012 <- net %>%
  filter(year == 2012) %>%
  mutate(yday = yday(date)) %>%
  dplyr::select(site, yday) %>%
  left_join(site_data) %>%
  ungroup() %>%
  dplyr::select(site, elev.mean, lat, lon, yday) %>%
  unique() %>%
  arrange(site, yday) %>%
  unite(site.day, site, yday, sep = "_", remove = "FALSE") %>%
  dplyr::select(-site)

bb_sitepair.2012 <- formatsitepair(BB_bioData.2012, bioFormat = 1, siteColumn = "site.day", dist = "jaccard",
                              XColumn = "lon", YColumn = "lat", predData = BB_predData.2012)

### Call gdm()
gdm_bb.2012 <- gdm(bb_sitepair.2012, geo = TRUE)
gdm.varImp(bb_sitepair.2012, geo = TRUE)

summary(gdm_bb.2012)
plot(gdm_bb.2012)
plotUncertainty(bb_sitepair.2012, sampleSites = 0.7, bsIters = 50, geo  = TRUE)
```

### Now let's look at interaction beta diversity. I don't think p-values would add anything to the interpretation I get from the plotted uncertainty, so it's okay that I can't get gdm.varImp() to work.
```{r}
### 2010
sitepair_2010 <- beta_2010 %>%
  ungroup() %>%
  filter(metric == "WN") %>%
  mutate(weights = rep(1, n())) %>%
  dplyr::select(distance = value, weights,
         s1.xCoord = lon1, s1.yCoord = lat1, 
         s2.xCoord = lon2, s2.yCoord = lat2, 
         s1.elev.mean = elev1, s2.elev.mean = elev2) %>%
  data.frame()
